---
title: "Basic R Exercise 2"
author: "Brianna Kincaid"
date: "February 2, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Matrix problems

### 1. Suppose

$$ A = \begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

###(a) Check that $A^3 = \mathbf{0}$  
###(b) Replace the third column of A by the sum of the second and third columns  

******   
###First, produce A

```{r}
A <- matrix(c(1,1,3,5,2,6,-2,-1,-3), nrow = 3, byrow = TRUE)
A
```

###Then, add the columns 2 and 3 and assign the sum to the third column

```{r}
A[,3] <- A[,2] + A[,3]

A
```


******

### 2. Create the following matrix B with 15 rows

$$B = \begin{bmatrix}
10 & -10 & 10\\
10 &  -10& 10\\
 \cdots & \cdots  & \cdots \\
10 & -10 & 10
\end{bmatrix} $$

###Calculate the 3x3 matrix $B^TB$.  You can make this calculation with the function crossprod().  See the documentaion. 

******
###First, produce B

```{r}
B <- matrix(c(10,-10,10),nrow=15, ncol=3, byrow=TRUE)
B
```

###Then, calculate the 3x3 matrix $B^TB$ using crossprod(). Given matrices x and y as arguments, the function crossprod() returns the matrix cross product. It is equivalent to doing matrix multiplication (%*%) of the transpose of x with y. 


```{r}
tB_B <- crossprod(B,B)
tB_B
```

******

### 3. Create a 6 x 6 matrix matE  with every element equal to 0.  check what the functions row() and col() return when applied to matE.Hence, create the 6 x 6 matix:

$$\begin{bmatrix}
0 &1  & 0 & 0 & 0 & 0\\
1 &0 & 1 &0  & 0 &0 \\
0 & 1 & 0 & 1 & 0 & 0\\
0 & 0 & 1 & 0 & 1 & 0\\
0& 0 & 0 & 1 & 0 & 1\\
0& 0 & 0 & 0 & 1 & 0
\end{bmatrix} $$  

******
###Here is matE, a 6x6 matrix of 0's followed by row(matE) and col(matE)
```{r}
matE <- matrix(rep(0,36), nrow = 6, byrow = TRUE)  
matE

# Note what the functions row() and col() do
row(matE) #returns a matrix of integerrs indicating their row number in a matrix-like object, or a factor indicating the row labels
col(matE) #Returns a matrix of integers indicating their column number in a matrix-like object, or a factor of column labels.

```


```{r}
# With a little experimentation you would see
# that the specified pattern is in the |1|'s
row(matE)-col(matE)
```

```{r}
# so you use the locations of the 1's to modify matE 
matE[abs(row(matE)-col(matE))==1] <- 1
matE
```

******


### 4.  Look at the help for the function outer().  Now, create the following patterned matrix:

$$\begin{bmatrix}
0 & 1 & 2 & 3 & 4\\
1 & 2 & 3 & 4 & 5\\
2 & 3 & 4 & 5 & 6\\
3 & 4 & 5 & 6 & 7\\
4 & 5 & 6 & 7 & 8
\end{bmatrix} $$

###From the documentation of outer product: The outer product of the arrays X and Y is the array A with dimension c(dim(X), dim(Y)) where element A[c(arrayindex.x, arrayindex.y)] = FUN(X[arrayindex.x], Y[arrayindex.y], ...). Can be used like 
###outer(X, Y, FUN = "*", ...)
###or
###X %o% Y



```{r}

a <- 0:4
A <- outer(a,a,"+")
A
```

###Use outer() a little more to make sure you get it.

```{r}
B <- outer(a,a, "*")
B

# and
b <- 5:10
C <- outer(a,b,"+")
C

# and finally  --  make sure you check the values.
D <- outer(b,a, "%%")
D
```


### 5.  Create the following patterned matrices.  Your solutions should be generalizable to enable creating larger matrices with the same structure.

(a) 

$$\begin{bmatrix}
0 & 1 & 2 & 3 & 4\\
1 &2 & 3 & 4 & 0\\
2 & 3 & 4 & 0 & 1\\
3 & 4 & 0 & 1 & 2\\
4 & 0 & 1 & 2 & 3
\end{bmatrix}  $$

******

```{r}
dim <- 4
a <- 0:dim
A <- outer(a,a,"+") 
A <- A - (A>(dim))*(dim+1)
A
```

******

(b)

$$ \begin{bmatrix}
0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9\\
1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 0\\
\vdots  & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
8 & 9 & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7\\
9 & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8
\end{bmatrix} $$

******

```{r}
dim <- 9
b <- 0:dim
B <- outer(b,b,"+") 
B <- B - (B>(dim))*(dim+1)
B
```
#this is the same matrix as a, with a larger dimension

******

(c)

$$\begin{bmatrix}
0 & 8 & 7 & 6 & 5 & 4 &3  & 2 & 1\\
1 & 0 & 8 & 7 & 6 & 5 & 4 & 3 & 2\\
2 & 1 & 0 & 8 & 7 & 6 & 5 & 4 & 3\\
3 & 2 & 1 & 0 & 8 & 7 & 6 & 5 & 4\\
4 & 3 & 2 & 1 & 0 & 8 & 7 & 6 & 5\\
5 & 4 & 3 & 2 & 1 & 0 & 8 & 7 & 6\\
6 & 5 & 4 & 3 & 2 & 1 & 0 & 8 & 7\\
7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 & 8\\
8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0
\end{bmatrix} $$

******

```{r}
dim <- 8
c1 <- 0:dim
c2 <- c(0,dim:1)
C <- outer(c1,c2,"+") 
C <- C - (C>(dim))*(dim+1)
C
```

******

### 6.  Solve the following system of linear equations by setting up and solving the matrix equation Ax = y.

### Make use of the special form of the matrix. A. The method used for the solution easily generalise to a larger set of equations where the matrix A has the same structure; hence the solution should not involve typing in every number of A.   


$x_1 + 2x_2 + 3x_3 + 4x_4 + 5x_5 =  7$  
$2x_1 +  x_2 + 2x_3 + 3x_4 + 4x_5 = -1$  
$3x_1 + 2x_2 +  x_3 + 2x_4 + 3x_5 = -3$  
$4x_1 + 3x_2 + 2x_3 +  x_4 + 2x_5 =  5$  
$5x_1 + 4x_2 + 3x_3 + 2x_4 +  x_5 = 17$  

******
```{r}
dim <- 5
A <- array(0,dim=c(dim,dim))

A <- row(A)-col(A)+1
A <- (A<1)*2 + abs(A)
A

y = c(7,-1,-3,5,17)
x <- solve(A,y)
x

```
******
 
### 7. Create a 6 x 10 matrix of random integers chosen from 1,2,...,10 by executing the following two lines of code:

```{r}
set.seed(75)  
aMat <- matrix(sample(10, size=60, replace=TRUE), nr=6) 
``` 

###Use the matrix you have created to answer these questions:  

###(a) Find the number of entries in each row which are greater than 4.

******
```{r}
rowSums(aMat>4)
```
******

###(b) Which rows contain exactly two occurrences of the number seven?

```{r}
which(rowSums(aMat==7)==2)
```

###(c) Find those pairs of columns whose total (over both columns) is greater tnan 75.  The answer should be a matrix with two columns; so, for example, the row (1,2) in the output matrix means that the sum of columns 1 and 2 in the original matrix is greater than 75.  Repeating a column is permitted; so, for exalmple, the final output matrix could contain the rows (1,2), (2,1), and (2,2).  

```{r}
num_cols <- dim(aMat)[2]
i <- 1:num_cols

library(gtools)
all_indeces <- permutations(n=10,r=2,v=i,repeats.allowed=T)

num_perm <- nrow(all_indeces)

count <- 0
indeces <- matrix(c(0,0),nrow=num_perm,ncol=2)
for (i in 1:num_perm) {
  if (sum(aMat[,all_indeces[i,1]]) + sum(aMat[,all_indeces[i,2]]) > 75) {
    count <- count+1
    indeces[count,] <- all_indeces[i,]
  }

}

indeces <- indeces[1:count,]
indeces
```

###What if repetitions are not permitted? Then only (1,2) from (1,2),(2,1) and (2,2) would be permitted.

```{r}
num_cols <- dim(aMat)[2]
i <- 1:num_cols

library(gtools)
all_indeces <- permutations(n=10,r=2,v=i,repeats.allowed=F)

num_perm <- nrow(all_indeces)

count <- 0
indeces <- matrix(c(0,0),nrow=num_perm,ncol=2)
for (i in 1:num_perm) {
  if (sum(aMat[,all_indeces[i,1]]) + sum(aMat[,all_indeces[i,2]]) > 75) {
    count <- count+1
    indeces[count,] <- all_indeces[i,]
  }

}

indeces <- indeces[1:count,]
indeces
```

### 8. Calculate

###(a) $\sum\limits_{i=1}^{20} \sum\limits_{j=1}^5 \frac{i^4}{(3+j)}$

******
```{r}

sum((1:20)^4) * sum(1/(3+(1:5)))

# or

sum(outer((1:20)^4, (3+(1:5)), "/"))

```
******

###(b) $\sum\limits_{i=1}^{20} \sum\limits_{j=1}^5 \frac{i^4}{(3+ij)}$

******
```{r}

for (i in 1:20){
  for (j in 1:5){
    x[i]<-(i^4)/(3+(i*j))
  }
  
}

result <- sum(x)
result

```
******


###(c) $\sum\limits_{i=1}^{10} \sum\limits_{j=1}^i \frac{i^4}{(3+ij)}$

******
```{r}

for (i in 1:10){
  for (j in 1:i){
    x[i]<-(i^4)/(3+(i*j))
  }
  
}

result <- sum(x)
result

```
******