---
title: "R Function Exercises"
author: "Brianna Kincaid"
date: "February 2, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is **\texttt{tmpFn1}**:
```{r}
tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}
```

Here is a simple example of **\texttt{tmpFn1}**:
```{r}
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b
```

Here is **\texttt{tmpFn2}**:
```{r}
tmpFn2 <- function(xVec2){
  n = length(xVec2)
  return(xVec2^(1:n)/(1:n))
}
```

Here is a simple example of **\texttt{tmpFn2}**:
```{r}
c <- tmpFn2(a)
c
```

******

(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$

******

Here is the function **\texttt{tmpFn3}**:
```{r}
tmpFn3 <- function(x,n){
  exp <- 1:n
  return(1+sum((x^exp)/exp))
}
```
\newpage
Here is a simple example of **\texttt{tmpFn3}** when $x=5$ and $n=2$. The expected result is $18.5$:
```{r}
b <- tmpFn3(5,2)
b
```
******

2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$

******
```{r}
tmpFn <- function(xVec){
  n = 3:length(xVec)
  return((xVec[n-2]+xVec[n-1]+xVec[n])/3)
}
```

Here is the function in use with the following input: \texttt{tmpFn}(c(1:5,6:1))
```{r}
t <- tmpFn(c(1:5,6:1))
t
```
******

3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}. Hence plot the function $f(x)$ for $-3 < x <3$.

******
Here is the function:
```{r}
tmpFn <- function(xVec){
  result <- rep(0,length(xVec))
  for(i in 1:length(xVec)){
  if (xVec[i] < 0) {result[i]<-(xVec[i]^2 + 2*xVec[i] + 3)}
  if (xVec[i] >= 0 && xVec[i] <2) {result[i] <- (xVec[i]+3)}
  if (xVec[i] >= 2) {result[i] <- (xVec[i]^2 + 4*xVec[i] - 7)}
  }
  return(result)
}
```
Here is the plot for $-3 < x < 3$. 
```{r}
x <- seq(-3,3,0.1)
y <- tmpFn(x)

plot(x,y,ylab='tmpFn(x)',main='tmpFn for -3 < x < 3')
```

******

4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  

******

Here is the function:
```{r}
matFn <- function(A) {
  B <- (A%%2==1)*2
  B[B==0]<-1
  return(A*B)
}
```

Here is the result of using the function on the matrix
$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

The expected result is:
$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$
```{r}
A <- matrix(c(1,1,3,5,2,6,-2,-1,-3),nrow=3,ncol=3,byrow=TRUE)
B <- matFn(A)
B
```
******

5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:

$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$

******
Here is the function:
```{r}
matFn2 <- function(n,k) {
  A <- diag(n)*k
  B <- abs(row(A)-col(A))
  B[B!=1]<-0
  return(B+A)
}
```
Here is a special case, when $n=5$ and $k=2$:
```{r}
C <- matFn2(5,2)
C
```
******

6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   

Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.
\newpage

******
```{r}
quadrant <- function(alpha) {
  
  if(alpha > 360){alpha <- alpha - 360*(as.integer(alpha/360))}
  
  if(alpha >= 0 & alpha < 90) {quad <- 1}
  if(alpha >= 90 & alpha < 180) {quad <- 2}
  if(alpha >= 180 & alpha < 270) {quad <- 3}
  if(alpha >= 270 & alpha < 360) {quad <- 4}
  
  return(quad)
}
```
******

### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

* $k$ = the day of the month  
* $y$ = the year in the century  
* $c$ = the first 2 digits of the year (the century number)  
* $m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   

For example, the date 21/07/1963 has $m = 5, k = 21, c = 19, y = 63$; the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  

Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year. Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.

******
```{r}

weekday <- function(day,month,year){
  days = c('Sunday','Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')
  #setting the values
  if(month==2 | month==1) {m <- (month-2)+12} else {m <- (month-2)}
  k <- day
  y <- year
  while(y>100){y <- y-100}
  c <- (year-y)/100
  #Zeller's congruence 
  f <- (as.integer((2.6*m) -0.2) + k + y + as.integer(y/4) + as.integer(c/4) - (2*c))%%7
  return(days[f])
}

```
******

(b) Does your function work if the input parameters \texttt{day, month} and \texttt{year} are vectors with the same length and valid entries?

******
```{r}
weekday <- function(day,month,year){
  days = c('Sunday','Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')
  m <-0
  cent<-0
  k<-0
  y<-0
  f <- 0
  
  for(i in 1:length(day)){
  #setting the values
  if(month[i]==2 | month[i]==1) {m[i] <- (month[i]-2)+12} else {m[i] <- (month[i]-2)}
  k[i] <- day[i]
  y[i] <- year[i]
  while(y[i]>100){y[i] <- y[i]-100}
  cent[i] <- (year[i]-y[i])/100
  
  #Zeller's congruence 
  f[i] <- (as.integer((2.6*m[i]) -0.2) + k[i] + y[i] + as.integer(y[i]/4) + as.integer(cent[i]/4) - (2*cent[i]))%%7
  }
  
  return(days[f])
}

```

Let's test the function for a input parameters that are vectors. 
```{r}
day <- c(4,5,1,23,15,8)
month <- c(12,11,4,5,7,2)
year <- c(1996,1987,1972,2018,2014,2017)

vector_result <- weekday(day,month,year)
vector_result

```
******

###8.

(a) Supposed $x_0=1$ and $x_1=2$ and 
$$x_j = x_{j-1} + \frac{2}{x_{j-1}}$$ for $$j=1,2, ....$$
Write a function \texttt{testLoop} which takes the single argument $n$ and returns the first $n-1$ values of the sequence $\left \{x_j  \right \}_{j\geq 0}$: that means the values of $x_0,x_1,x_2,...,x_{n-2}$.

******

```{r}
testloop <- function(n) {
  x <- rep(0,n-1)
  x[1] <- 1 #x0
  x[2] <- 2 #x1
  for (i in 3:(n-1)) {x[i] = x[i-1] + 2/(x[i-1])}
  return(x)
  
}
```
******
(b) Now write a function \texttt{testLoop2} which takes a single argument \texttt{yVec} which is a vector. The function should return

$$\sum_{j=1}^{n}e^{j}$$

where $n$ is the length of \texttt{yVec}.

******

```{r}
testloop2 <- function(yVec) {
  n <- lenght(yVec)
  j <- 1:n
  return(sum(e^j))
  
}
```
******

9.*Solution of the difference equation $x_n = rx_{n-1}(1-x_{n-1})$, with starting value $x_1$.*

(a) Write a function  \texttt{quadmap(start,rho,niter)} which returns the vector $(x_1,...,x_n)$ where $x_k = rx_{k-1}(1-x_{k-1})$ and
*\texttt{niter} denotes $n$,
*\texttt{start} denotes $x_1$, and
*\texttt{rho} denotes $r$.

******
Here is the function:
```{r}
quadmap <- function(start, rho, niter) {
  x <- rep(0,niter)
  x[1] <- start
  
  for(i in 2:niter) {
  x[i] <- rho*(x[i-1])*(1-x[i-1])
  }
  
  #for some reason this worked when I put it in a for loop but it did not work when I 
  #pulled it out of the for loop -- can't really figure out why
  
  return(x)
}

```
Now let's try out the function for $r=2$ and $0 < x1 < 1$. $x_n$ should go to $0.5$ as $n \to \infty$.
```{r}
y <- quadmap(.1,2,20)
n <- 1:20
plot(n,y,type="l",ylab='quadmap',main='quadmap() for r=2 and 0 < x1 < 1')
```
As you can see from the plot, the values of x go to 0.5 as n gets large.

Now let's try out the function with $start=0.95$, $rho=2.99$, and $niter=500$.
```{r}
tmp <- quadmap(start=0.95, rho=2.99,niter=500)
plot(tmp,type="l")
plot(tmp[300:500], type="l")

```

******

(b) Now write a function which determines the number of iterations needed to get $\left | x_n - x_{n-1} \right |< 0.02$. So this function has only 2 arguments: \texttt{start} and \texttt{rho}. (For \texttt{start=0.95} and \texttt{rho=2.99}, the answer is 84.)

******

```{r}
iterations <- function(start,rho) {
  x <- start
  x[2] <- rho*(start)*(1-start)
  n <- 2 #minimum value that n must be 
  
  while(abs(x[n]-x[n-1])>= 0.02) {
    n <- n+1
    x[n] <- rho*(x[n-1])*(1-x[n-1])
  }
  
  return(n)
}

```
Here is the function called with the inputs \texttt{start=0.95} and \texttt{rho=2.99}.
```{r}
iterations(0.95,2.99)
```

******

10.

(a) Given a vector $(x_1,...,x_n)$, the sample autocorrelation of lag $k$ is defined to be 
$$r_k = \frac{\sum_{i=k+1}^{n}(x_i - \bar{x})(x_{i-k}-\bar{x})}{\sum_{i=1}^{n}(x_i - \bar{x})^2}$$
Write a function \texttt{tmpFn(xVec)} which takes a single argument \texttt{xVec} which is a vector and returns a \texttt{list} of two values: $r_1$ and $r_2$.
In particular, find $r_1$ and $r_2$ for the vector (2,5,8,...,53,56).

******
Here is the function.
```{r}
tmpFn <- function(xVec){
  xb <- mean(xVec) #x-bar
  
  ## r1
  it <- 2:length(xVec)
  id <- 1:length(xVec)
  r1 <- sum((xVec[it]-xb)*(xVec[it-1]-xb))/sum(((xVec[id]-xb)^2))
  
  ## r2
  it <- 3:length(xVec)
  id <- 1:length(xVec)
  r2 <- sum((xVec[it]-xb)*(xVec[it-2]-xb))/sum(((xVec[id]-xb)^2))
  
  return(c(r1,r2))
}

```

Now let's test the function for the vector $(2,5,8,...,53,56)$.
```{r}
input <- seq(2,56,3)
output <- tmpFn(input)

r1 <- output[1]
r1
r2 <- output[2]
r2
```
******

(b) Generalize the function so that it takes two arguments: the vector \texttt{xVec} and an integer \texttt{k} which lies between 1 and $n-1$ where $n$ is the length of \texttt{xVec}. The function should return a vector of the values $(r_0=1,r_1,...,r_k)$.

******

```{r}
tmpFn2 <- function(xVec,k) {
  r <- 1

  #r[2]=r1, r[3]=r2,... r[k+1]=rk
  
  for(i in 2:(k+1)){
    r[i] <- sum((xVec[i:length(xVec)]-mean(xVec))*(xVec[(i:length(xVec))-(i-1)]-mean(xVec)))/sum(((xVec[1:length(xVec)]-mean(xVec))^2))
  }
  
  return(r)
  
}
```




